Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-03 doc/sc2012/soop.tex
full draft of SOOP

\input{mapping_fig}

@@

\begin{figure}
\includegraphics[scale=0.48]{figs/CircuitMem.pdf}
\caption{Tasks and data for the circuit simulation on a cluster of GPUs.}
\end{figure}

@@

>>>>

After a processor $p$ is chosen for a task, $t$ executes entirely on $p$.
Before execution, however, we must {\em map} $t$, i.e., choose the
physical instances for $t$'s logical regions.  A physical instance is assigned
to a particular memory in the machine and holds a snapshot of the region's data
from some point in time; a {\em valid} physical instance has data that reflects
all the updates of all previously mapped tasks.

The region tree used to calculate logical dependences for $t$ is also
used to hold mapping information.  Each logical region $r$ maintains a
list of its physical instances and which sibling tasks are using which
instances and with what permissions.  The important correctness
property established by the dependence analysis phase is that all
tasks $t$ depends on map before $t$ maps and no task that depends on
$t$ maps before $t$ finishes mapping.  Thus, the region tree at the
moment $t$ maps contains a consistent view of what the state of memory
will be after $t$'s dependence predecessors have executed.

There are two steps to mapping a particular region argument $r$ of
task $t$.  First, the mapper must ensure there is at least one {\em valid}
physical instance of $r$.  For example, assume there is one physical instance $s$ of
logical region $r$ and that a task $t'$ that mapped prior to $t$ is recorded as writing a
physical instance $s_i$ of subregion $r_i$ of $r$.  At this point there
is no valid instance of $r$, because $s$ does not include
the updates to $s_i$.  To restore $s$ to a valid instance of $r$ the
system must shedule an operation after $t'$ to copy
$s_i$ back into $s$.

In the second step, the mapping interface is used to select one of the
valid physical instances of $r$ or to create a new one.  Typical
policies are to to choose the existing physical instance that is
closest to $t$'s processor, or to create a new instance (by issuing a
copy operation) in the memory closest to $t$'s processor.
We focus here on the first step, establishing a valid physical instance.
The mapping interface is discussed further in Section~\ref{sec:mapping}.

Detecting or creating a valid instance is, at its core, a dependence analysis on
physical instances and thus similar to the mapping dependence analysis on logical
regions already described in Section~\ref{sec:dep}.  Consider any task $t'$ using
physical instance $s'$ of logical region $r'$.  There are four cases.
\begin{enumerate}
\item If $t'$ has only read permission
for $r'$ it cannot invalidate any instance of region $r$ and nothing needs to be done.

\item If $t'$ has reduce permission for $r'$ and $t$ has reduce
  permission for $r$ and both use the same reduction operator, then
  nothing needs to be done.  This is the only case in which a writer
  does not require a valid instance---because reductions can be
  reordered, the recombining of the instances of $r$ and $r'$ (if they
  overlap) can be deferred to a later consumer of the data (see case
  4).

\item If $t'$ has write permission for $r'$ and $r$ and $r'$ may
  overlap (i.e., $\lca{r}{r'}$ is either a region or an aliased
  partition), then $s'$ must be copied to an instance $s''$ of
  $\lca{r}{r'}$ and then a fresh instance of $r'$ created from $s''$
  (if either $r = \lca{r}{r'}$ or $r' = \lca{r}{r'}$ one of the two
  copy operations is omitted).
 IS s' REMOVED FROM THE TREE?

\item If $t'$ has reduce permission for $r'$, $t$ has read or write
  permission for $r$, and $r$ and $r'$ may overlap, then $s'$ must be
  reduced (using $t'$'s reduction operator) to an instance $s''$ of
  $\lca{r}{r'}$ and then a fresh instance of $r'$ created from $s''$.
  (If $r' = \lca{r}{r'}$ the reduction is unneeded, and if $r =
  \lca{r}{r'}$ the copy is unneeded.)
 IS s' REMOVED FROM THE TREE?
\end{enumerate}

Not all of these cases can occur simultaneously for a particular
region argument $r$ of task $t$.  There may always be multiple
previous readers (case 1), but the mapping dependence analysis
guarantees that there would only be one previous writer (case 3) or
one or more reducers (cases 2 and 4).  To map $r$, we walk from $r$'s
root ancestor in the region forest to $r$, along the way exploring
off-path subtrees to find region instances satisfying cases 3 and
4; the details are similar to the implementation of dependence
analysis in Section~\ref{sec:dep}.

As part of the walk any required copy and reduction operations are
issued to construct a valid instance of $r$.  These operations are
deferred, waiting to execute on the tasks that produce the data they
copy or reduce. Similarly, the start of $t$'s execution is made
dependent on completion of the copies/reductions that construct the
instance of $r$ it will use.

%
%possibly make this consistent with the way we actually map circuit?
%
Consider again the four example tasks from Listing~\ref{lst:code_ex}.
We focus only on the mapping of the node regions of the region tree in
Figure~\ref{sfig:part_fig:tree}.  For this example, we assume a simple 
machine with $n$ processors each with a local memory, and we map the $i$th
piece of the circuit to the $i$th processor's memory.  Thus, all the
region arguments of 
{\tt cal\_new\_currents(pieces[0])}
and
{\tt cal\_new\_currents(pieces[1])}
are mapped to instances in the memory's of processor's 0 and 1, respectively.  The tasks themselves
are also mapped to processors 0 and 1, respectively.

When {\tt distribute\_charge(pieces[0])} executes, it uses the same
instances on processor 0 and the task also runs on processor 0.
Enlarging the example to include more tasks, note that the region
arguments of {\tt distribute\_charge(pieces[i])} overlap with {\tt
  pieces[0]} both because the shared and ghost partitions are of the
same data and because te ghost partitions are not disjoint.  However,
{\tt distribute\_charge} only performs commutative reductions on all
of these regions and so the subregion instances can be used as is on
each of the processors for all {\tt distribute\_charge} tasks.
Finally, {\tt update\_voltages(pieces[0])} reads and writes {\tt
  pieces[0].rn\_shr}.  To guarantee the task sees the correct data,
all of the shared and ghost partitions must be reduced back to the
region of all shared nodes {\tt p\_nodes\_pvs[1]} (see
Figure~\ref{sfig:part_fig:tree}, after which a fresh instance of {\tt
  pieces[0].rn\_shr} can be copied from {\tt p\_nodes\_pvs[1]} to node
0.

%\input{mapping_fig}

Figure~\ref{fig:gpumapping} shows a more involved mapping of the circuit simulation
on a cluster of GPUs.  The $i$th private node and wire regions are mapped to the
$i$th GPU's DRAM and simply left there for all three phases of the simulation.  The $i$th
shared and ghost regions are mapped to zero-copy memory most of the time, but have to be
reduced out to the global (GASNet) memory and a new instances of the  shared  regions
created again with the updated information before the {\tt update\_voltage} tasks execute.
A similar situation arises before {\tt calc\_new\_currents} tasks run again: the updates
to the shared nodes by {\tt update\_voltage} have to be propagated to the ghost nodes before
continuing.

\begin{figure}
\includegraphics[scale=0.48]{figs/CircuitMem.pdf}
\caption{Tasks and data for the circuit simulation on a cluster of GPUs.}
\label{fig:gpumapping}
\end{figure}

@@

Once a task $t$ has been mapped it enters the execution stage of the SOOP.  When all of
the operations (other tasks and copies) on which $t$ depends have completed, $t$ is
launched on the processor.  When $t$ launches subtasks, they are recursively scheduled as siblings by
the SOOP using $t$ as the parent task and $t$'s region arguments as the roots of the region forest.

@@

Once a task is done executing the various pieces of state associated with it can be reclaimed.
Dependence and mapping information can be removed from the region tree; in point of fact, 
as an optimization we are often able to remove tasks from the region tree well before the task terminates.
The most involved aspect of clean-up is colleting physical instances that are no longer in use.
We use a distributed reference counting scheme to track the number of tasks using physical instances;
when this number goes to 0 the physical intance can be reclaimed.
